\large{\bf Looking forwards and backwards through locally regulated populations }

### 2023-03-20 forwards-and-backwards.tex
cites
edits with alison

	\comment{Need to thank KAVLI, Paris Cit\'e, various funders}

@@

@@

By replacing $\rho$ by
$\rho^\epsilon(\cdot)=\rho(\cdot/\epsilon)/\epsilon^d$, 
we can also scale the ``width'' of the 
region over which we measure local population density. 
When the population follows~(\ref{general deterministic limit}), 
we expect that if we take a second limit of $\epsilon \to 0$,
thus scaling the kernels appearing in $r$, $\gamma$, and $F$ and making interactions are pointwise,
we should recover a nonlinear PDE.
We verify that this is indeed the case in two important examples:
a special case of the porous medium equation with a logistic growth term, in which we 
retain a signal of our two stage approach to modelling in the nonlinear diffusion term
and the limiting equation takes the form 
\begin{equation}
	\label{PME1}
	\partial_t \varphi = \Delta (\varphi^2)+\varphi(1-\varphi);
\end{equation}
and a wide class of semi-linear PDEs of the form 
\begin{equation}
\label{semilinear PDE}
\partial_t \varphi = \DG^*\varphi+ \varphi F(\varphi),
\end{equation}
which includes the Fisher-KPP equation and the Allen-Cahn equation.

@@

We remark that such results on (one-step) convergence to reaction diffusion 
equation limits have been
achieved for a variety of interacting particle systems. The vast majority of these
results concern spin systems \comment{REFs}. However,~\cite{flandoli/huang:2021} 
obtain the Fisher-KPP equation
as a scaling limit of interacting diffusions in
$\IR^d$. In that setting a major challenge (also apparent in our models), not
faced in spin systems, is 
the control of concentration of particles. We also note that the scaling 
considered in~\cite{flandoli/huang:2021} is different from that considered
below. In their language we consider only `moderate interactions', meaning that
the number of individuals in the neighbourhood over which we measure local
population density tends to infinity, whereas they also consider the situation in which
that number remains finite; furthermore.
On the other hand, we shall consider a scaling in which 
the reproduction mechanism is near-critical and look at large timescales,
while \citet{flandoli/huang:2021} only scale the number of
particles in the system and the size of the neighbourhood over which indivduals interact.

@@

The shape of the wave is not determined solely by the reaction term.
For example, as a result of the nonlinear diffusion, 
for suitable initial conditions, the solution to the one-dimensional 
porous medium equation with logistic growth~(\ref{PME1}) converges to a travelling
wave with a sharp cut-off; i.e., in contrast to the classical 
Fisher KPP equation, the solution at time $t$
vanishes beyond $x=x_0+ct$ for some constant wavespeed $c>0$, \citep{kamin/rosenau:2004}.
As a first step towards understanding what we should expect in models with
nonlinear diffusion, one can ask about the position of an ancestral lineage
relative to the wavefront in the deterministic models. In 
Section~\ref{ancestral lineages for travelling waves}
we shall see that in our framework, even with logistic growth, the nonlinear diffusion 
corresponding to the porous medium equation results
in a stationary distribution that is concentrated behind the wavefront, leading 
us to conjecture that in the stochastic equation
the cooperative behaviour captured by the nonlinear diffusion will also 
result in a qualitatively different pattern of coalescence to that seen under the 
stochastic Fisher-KPP equation. Indeed, we believe that it should be feasible to 
show that in an appropriate limit one recovers a Kingman coalescent.

@@

We shall write $\mu_\theta(x, X/N)$ for the per-capita death rate of mature individuals
in the population.
In order for the 
population density to change over timescales of order $\theta$, we should like the net 
per capita reproductive rate 
to scale as $1/\theta$. 
In classical models, in which $r$, $\gamma$, and $\mu$ are constant, 
this quantity is simply $r\gamma -\mu$. 
Here, because production of juveniles and their establishment are
mediated by population density measured relative to different points, 
the net reproductive rate will take a more complicated form. 
In particular, the total rate of production of mature offspring by an individual at $x$ will be
\[
	\gamma\big(x,\smooth{\gamma}X(x)/N\big)\int r\big(y, \smooth{r}X(y)/N\big)q_{\theta}(x,dy).
\]
Nonetheless it will be convenient to define the death rate $\mu_\theta$ in terms of 
its deviation from $r\gamma$. 
To this end, we define
the death rate of an individual at $x$ to be
\begin{align} \label{eqn:mu_defn}
    \mu_\theta(x)
    =
    \max\left\{0, 
        r(x, \smooth{r} X(x) / N) \gamma(x, \smooth{\gamma} X(x) / N)
        - \frac{1}{\theta} F(x, \smooth{F} X(x) / N)
    \right\} ,
\end{align}
where this equation defines $F(x, m) : \IR^d \times \IR_{\ge 0} \to \IR$,
the net per capita reproductive rate, scaled by $\theta$,
and $\rho_F$ is again a smooth kernel. 
It will be convenient to work with $F$ instead of $\mu$.

@@

\comment{Do we need a "Terminology" heading for this?}
\begin{terminology*}
In our prelimiting model,
the population is represented by a point measure in which each
individual is assigned a mass $1/N$. We use the term ``population density'' for
this process, as it is supposed to measure population size relative to a 
nominal occupancy of $N$ individuals per unit area. There is no implication 
that the measure representing the population is absolutely continuous with
respect to Lebesgue measure; indeed in the prelimit it is certainly not.
\end{terminology*}

@@

It is convenient to characterise the process as a solution to a martingale problem.
We write $C_b^\infty(\IR^d)$ for the space of bounded smooth functions on $\IR^d$, and, where
convenient, we
write $\langle f, \eta \rangle = \int_{\IR^d} f(x) \eta(dx)$.
\begin{definition}[Martingale Problem Characterisation]
    \label{defn:mgale_construction}
	For each value of $N$ and $\theta$, and each purely atomic 
$\eta_0^N\in\measures$ with atoms of mass $1/N$,
$(\eta^N_t)_{t \geq 0}$ is the (scaled) empirical measure of a birth-death 
process with c\`adl\`ag paths in $\measures$ for which,
for all $f \in C^{\infty}_{b}(\IR^d)$,
writing $q_\theta(x, dy)$ for the Gaussian kernel
with mean $x + \meanq(x)/\theta$ and covariance $\covq(x) / \theta$,
\begin{equation}
    \label{eqn:eta_martingale}
\begin{aligned}
M^N_t(f)
&:=  \langle f, \eta^N_t \rangle
        -\langle f, \eta^N_0 \rangle
 \\ &\qquad {}
 -  \int_{0}^{t}\bigg\{
        \int\left( \int \theta
             \left(
                f(z)r(z, \eta^N_{s})
                - f(x)r(x, \eta^N_{s})
            \right)
        q_\theta(x,dz) \right)
%\\ & \qquad \qquad \qquad \qquad {}
         \gamma(x, \eta^N_{s}) \eta^N_{s}(dx)
\\ & \qquad \qquad \qquad \qquad \qquad {}
    + \int f(x) F(x, \eta^N_{s}) \eta^N_{s}(dx)
    \bigg\} ds
\end{aligned}    
\end{equation}
is a martingale (with respect to the natural filtration), 
with quadratic variation
    \begin{equation} \label{eqn:prelimit_martingale_variation}
\begin{aligned} \relax
[ M^N(f)]_t =& 
\frac{\theta}{N} \int_{0}^{t}\bigg\{
    \Big\langle \gamma(x, \eta^N_{s})
        \int f^2(z)r(z, \eta^N_{s}) q_\theta(x,dz) 
    , \eta^N_{s}(dx) \Big\rangle \\
& \qquad \qquad {}
    + \Big\langle \mu_\theta(x, \eta^N_{s}) f^2(x) 
    , \eta^N_{s}(dx)\Big\rangle
    \bigg\}ds. 
\end{aligned}    
\end{equation}
\end{definition}
The form of~(\ref{eqn:eta_martingale}) and~(\ref{eqn:prelimit_martingale_variation})
is explained in Section~\ref{sec:heuristics}.
Note that since (juvenile) individuals are produced at rate $N \gamma \eta$, but each
has mass $1/N$,
these factors of $N$ cancel in~\eqref{eqn:eta_martingale}. 
Under our scaling, $N$ and $\theta=\theta(N)$ will tend to infinity in such a 
way that $\alpha:=\lim_{N\to\infty}\theta(N)/N$ exists and is finite.
From the expression~(\ref{eqn:prelimit_martingale_variation}) it is easy to guess that 
whether the limiting processes will be deterministic or stochastic
is determined by whether $\alpha$ is zero or nonzero.

@@

We give two more concrete conditions under which the fourth condition is satisfied.

@@

The purpose of the conditions that we have placed on the reproduction parameters is to 
ensure that the net per capita reproduction rate (before time scaling)
is order $1/\theta$. As remarked above, 
because of the non-local reproduction mechanism, it no longer suffices to assume that 
$r(x,\smooth{r}\eta(x))\gamma(x, \smooth{\gamma}\eta(x))-\mu_\theta(x)$ is of order $1/\theta$.
An instructive example is the case where $\gamma \equiv 1$ and $F \equiv 0$, so that $\mu_\theta = r$,
and the population has all individuals at a single location (so, $\eta = \delta_x$),
in which case the mean rate of change of the total population size
is $\int (r(y, N \smooth{r}(y) ) - r(x, N \smooth{r}(x))) q_\theta(x, dy)$;
the first condition of Lemma~\ref{lem:conditions on r} would ensure this is of order $1/\theta$.

If $r(x,m)$ is independent of $m$, then the conditions are easy to 
satisfy; they just require some regularity of $r$ as a function of $x$. 
Condition~\ref{control through r} is satisfied if 
$\|\nabla\rho_r\|\leq C\rho_r$ 
and $m\partial_mr(x,m)$, $m^2\partial_{mm}r(x,m)$ are bounded.
This will be satsified, for example, if $\rho_r$ decays 
exponentially.
On the other hand, it might, for example, seem more natural to take $\rho_r$ to be a Gaussian 
density with parameter $\sigma_r$, say. Then, 
as we check in Lemma~\ref{old 6.1}, \comment{this is referring to the
old Lemma 6.1, now to go in Appendix,}
Condition~\ref{control through gamma} is satisfied if $\rho_\gamma$ is also
Gaussian with parameter $\sigma_\gamma$ and $\sigma_\gamma>\sigma_r$. For large enough
$\theta$, this condition guarantees that 
$\sigma_r+1/\theta <\sigma_\gamma$, so that the establishment probability of a juvenile
is controlled by individuals that are already `felt' by the fecundity kernel $\rho_\gamma$
at the location of their parent.

@@

Our main results depend on two dichotomies:
Is the limiting process deterministic or a (generalized) superprocess?
And, are interactions pointwise in the limit or nonlocal?
Below we have results for deterministic limits with pointwise and nonlocal interactions,
and for superprocess limits with nonlocal interactions.
See Figure \ref{fig:super_vs_det_2d} for illustrative simulations,
from direct simulation of the process using SLiM \citep{haller_slim_2019}.

\begin{figure}
    \begin{center}
        \includegraphics{figures/ex1a/fkpp_123.locations}
        \includegraphics{figures/ex1b/fkpp_123.locations}
    \end{center}
    \caption{
        Snapshots of two simulations, with small $\alpha=\theta/N$ (left) and large 
	$\alpha =\theta/N$ (right).
        Simulations are run with a Fisher-KPP-like parameterization:
        birth and establishment are constant, while death increases linearly with density,
        at slope $1/\theta$.
        Left: $\alpha=0.1$. Right: $\alpha=10$.
        Other parameters were the same:
        dispersal and interactions distance were set to 1,
        and the equilibrium density is 10 individuals per unit area.
        \label{fig:super_vs_det_2d}
    }
\end{figure}

\paragraph{Scaling limits with nonlocal interactions:}

Recall that the process $(\eta_t^N)_{t\geq 0}$ takes its values in the space
${\cal D}([0,\infty),\measures)$ of c\`adl\`ag
paths on $\measures$. We endow $\measures$ with the topology of weak convergence
and 
${\cal D}([0,\infty),\measures)$ with the Skorohod topology.
A sequence of processes taking values in 
${\cal D}([0,\infty),\measures)$ is said to be tight if the corresponding
sequence of distributions
is tight, so that any infinite sequence has a weakly convergent subsequence.
Our first result establishes tightness of our rescaled population processes in 
the case in which interactions remain nonlocal under the scaling, and 
characterises limit points as solutions to a martingale problem.

\begin{theorem} \label{thm:nonlocal_convergence}
    Let $(\eta^N_t)_{t \geq 0}$
    be as defined in Definition~\ref{defn:mgale_construction}
	and assume that as $N \to \infty$, $\theta(N) \to \infty$
	in such a way that $\theta(N)/N \to \alpha$.
    (However, the kernels $\rho_r$, $\rho_\gamma$, and $\rho_F$
    remain fixed.) 
Suppose that Assumptions~\ref{def:model_setup} hold and, 
further, that $\{\eta_0^N\}_{N\geq 1}$ is a sequence of 
purely atomic measures, with $\eta^N$ comprised of atoms of mass $1/N$, which is
tight 
in $\measures$ and such that there exists $\lambda>0$ for 
which $\langle e^{\lambda\|x\|}, \eta_0^N(dx)\rangle<C<\infty$
with $C$ independent of $N$.
    Then the sequence of processes
    $(\eta^N_t)_{t \geq 0}$ is tight, and for any limit point
    $(\eta_t)_{t \geq 0}$, for every $f \in C^\infty_b(\IR^d)$,
    \begin{align} \label{eqn:limiting_mgale_problem}
        \begin{split}
        M_t(f)
            &:=
            \langle f(x), \eta_t(dx) \rangle
            -
            \langle f(x), \eta_0(dx) \rangle
            \\ & \qquad
            -
            \int_0^t \big\langle
                \gamma(x, \eta_s)
                \mathcal{B}\left(
                    f(\cdot) r(\cdot, \eta_s)
                \right)(x)
                %\\ &\qquad \qquad \qquad {}
                +
                f(x)
                F(x, \eta_s),
                \eta_{s}(dx)
            \big\rangle ds
        \end{split}
    \end{align}
    is a martingale (with respect to the natural filtration),
    with quadratic variation
    \begin{align} \label{eqn:limiting_mgale_variation}
        [ M(f) ]_t
        =
        \alpha
        \int_0^t
        \big\langle
            %\left(
	    2\gamma\left( x, \eta_{s} \right)
            r\left(x, \eta_{s} \right)% + \mu\left(x, \eta_{s} \right)\right)
            f^2(x),
            \eta_{s} (dx)
        \big\rangle ds. 
    \end{align}
    Furthermore, if $\alpha = 0$ the limit is deterministic.
\end{theorem}

@@

\begin{corollary} \label{cor:superprocess_uniqueness}
    Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
    if the martingale problem
    defined by equations~\eqref{eqn:limiting_mgale_problem} 
and~\eqref{eqn:limiting_mgale_variation}
    has a unique solution,
    then $(\eta^N_t)_{t \ge 0}$ converges weakly
to that solution
    as $N \to \infty$.
\end{corollary}
When $\alpha>0$, the limit points can be thought of as interacting superprocesses. For
example, when $r$ and $\gamma$ are constant, and 
$F(x,\eta_s)=1-\smooth{F} \eta_s(x)$,
we recover the superprocess with nonlinear death rates corresponding to logistic growth.
We are not aware of a general result to determine when we will have uniqueness of 
solutions to the martingale problem of Theorem~\ref{thm:nonlocal_convergence}
when $\alpha>0$.
However, the Dawson--Girsanov transform tells us that we have uniqueness in this special 
case of the superprocess with nonlinear death rates, and 
Perkins stochastic calculus (and its adaptation to a 
lookdown setting) provides uniqueness for cases with interactions in the 
dispersal mechanism of
the superprocess. We refer 
to~\cite{dawson:1993, perkins:1992}, and \cite{donnelly/kurtz:1999} 
for approaches to showing that the martingale problem is well-posed in the case $\alpha>0$.

@@

\begin{theorem}[Convergence to a PDE]
    \label{thm:local_convergence}
\comment{Need to add conditions on initial condition}

>>>>

	\comment{Need to thank KAVLI, Paris Cit\'e, various funders}
    \comment{make notation for $\mathcal{D}$ consistent}

@@

\comment{TODO: cite \citet{lam2023introduction,cantrell2004spatial}}

@@

By replacing $\rho$ by
$\rho^\epsilon(\cdot)=\rho(\cdot/\epsilon)/\epsilon^d$, 
we can also scale the ``width'' of the 
region over which we measure local population density. 
When the population follows~(\ref{general deterministic limit}), 
we expect that if we take a second limit of $\epsilon \to 0$,
thus scaling the kernels appearing in $r$, $\gamma$, and $F$ and making interactions pointwise,
we should recover a nonlinear PDE.
We verify that this is indeed the case in two important examples:
a special case of the porous medium equation with a logistic growth term, in which we 
retain a signal of our two stage approach to modelling in the nonlinear diffusion term
and the limiting equation takes the form 
\begin{equation}
	\label{PME1}
	\partial_t \varphi = \Delta (\varphi^2)+\varphi(1-\varphi);
\end{equation}
and a wide class of semi-linear PDEs of the form 
\begin{equation}
\label{semilinear PDE}
\partial_t \varphi = \DG^*\varphi+ \varphi F(\varphi),
\end{equation}
which includes the Fisher-KPP equation and the Allen-Cahn equation.

@@

We remark that such results on (one-step) convergence to reaction diffusion 
equation limits have been
achieved for a variety of interacting particle systems. The vast majority of these
results concern spin systems \comment{REFs}. However,~\cite{flandoli/huang:2021} 
obtain the Fisher-KPP equation
as a scaling limit of interacting diffusions in
$\IR^d$. In that setting a major challenge (also apparent in our models), not
faced in spin systems, is 
the control of concentration of particles. We also note that the scaling 
considered in~\cite{flandoli/huang:2021} is different from that considered
below. In their language we consider only `moderate interactions', meaning that
the number of individuals in the neighbourhood over which we measure local
population density tends to infinity, whereas they also consider the situation in which
that number remains finite;
furthermore we shall consider a scaling in which 
the reproduction mechanism is near-critical and look at large timescales,
while \citet{flandoli/huang:2021} only scale the number of
particles in the system and the size of the neighbourhood over which indivduals interact.

@@

The shape of the wave is not determined solely by the reaction term.
For example, as a result of the nonlinear diffusion, 
for suitable initial conditions, the solution to the one-dimensional 
porous medium equation with logistic growth~(\ref{PME1}) converges to a travelling
wave with a sharp cut-off; i.e., in contrast to the classical 
Fisher KPP equation, the solution at time $t$
vanishes beyond $x=x_0+ct$ for some constant wavespeed $c>0$, \citep{kamin/rosenau:2004}.
As a first step towards understanding what we should expect in models with
nonlinear diffusion, one can ask about the position of an ancestral lineage
relative to the wavefront in the deterministic models. In 
Section~\ref{ancestral lineages for travelling waves}
we shall see that in our framework, even with logistic growth, the nonlinear diffusion 
corresponding to the porous medium equation results
in a stationary distribution for the ancestral lineage
that is concentrated behind the wavefront, leading 
us to conjecture that in the stochastic equation
the cooperative behaviour captured by the nonlinear diffusion will also 
result in a qualitatively different pattern of coalescence to that seen under the 
stochastic Fisher-KPP equation. Indeed, we believe that it should be feasible to 
show that in an appropriate limit one recovers a Kingman coalescent.

@@

We shall write $\mu_\theta(x, X/N)$ for the per-capita death rate of mature individuals
in the population.
In order for the 
population density to change over timescales of order $\theta$, we should like the net 
per capita reproductive rate 
to scale as $1/\theta$. 
In classical models, in which $r$, $\gamma$, and $\mu$ are constant, 
this quantity is simply $r\gamma -\mu$. 
Here, because production of juveniles and their establishment are
mediated by population density measured relative to different points, 
the net reproductive rate will take a more complicated form. 
In particular, the total rate of production of mature offspring by an individual at $x$ will be
\[
	\gamma\big(x,\smooth{\gamma}X(x)/N\big)\int r\big(y, \smooth{r}X(y)/N\big)q_{\theta}(x,dy).
\]
Nonetheless, it will be convenient to define the death rate $\mu_\theta$ in terms of 
its deviation from $r\gamma$. 
To this end, we define
the death rate of an individual at $x$ to be
\begin{align} \label{eqn:mu_defn}
    \mu_\theta(x)
    =
    \max\left\{0, 
        r(x, \smooth{r} X(x) / N) \gamma(x, \smooth{\gamma} X(x) / N)
        - \frac{1}{\theta} F(x, \smooth{F} X(x) / N)
    \right\} ,
\end{align}
where this equation defines $F(x, m) : \IR^d \times \IR_{\ge 0} \to \IR$,
and $\rho_F$ is again a smooth kernel. 
\comment{FIX UP THIS, REFER FORWARD}
The function $F$ is nearly
the net per capita reproductive rate, scaled by $\theta$,
and would be equal to it in a nonspatial model;
but 
It will be convenient to work with $F$ instead of $\mu$.

@@

\begin{terminology*}
In our prelimiting model,
the population is represented by a point measure in which each
individual is assigned a mass $1/N$. We use the term ``population density'' for
this process, as it is supposed to measure population size relative to a 
nominal occupancy of $N$ individuals per unit area. There is no implication 
that the measure representing the population is absolutely continuous with
respect to Lebesgue measure; indeed in the prelimit it is certainly not.
\end{terminology*}

@@

It is convenient to characterise the process as a solution to a martingale problem.
We write $C_b^\infty(\IR^d)$ for the space of bounded smooth functions on $\IR^d$, and, where
convenient, we
write $\langle f, \eta \rangle = \int_{\IR^d} f(x) \eta(dx)$.
\begin{definition}[Martingale Problem Characterisation]
    \label{defn:mgale_construction}
	For each value of $N$ and $\theta$, and each purely atomic 
$\eta_0^N\in\measures$ with atoms of mass $1/N$,
$(\eta^N_t)_{t \geq 0}$ is the (scaled) empirical measure of a birth-death 
process with c\`adl\`ag paths in $\measures$ for which,
for all $f \in C^{\infty}_{b}(\IR^d)$,
writing $q_\theta(x, dy)$ for the Gaussian kernel
with mean $x + \meanq(x)/\theta$ and covariance $\covq(x) / \theta$,
\begin{equation}
    \label{eqn:eta_martingale}
\begin{aligned}
M^N_t(f)
&:=  \langle f, \eta^N_t \rangle
        -\langle f, \eta^N_0 \rangle
 \\ &\qquad {}
 -  \int_{0}^{t}\bigg\{
        \bigg\langle \left( \int \theta
             \left(
                f(z)r(z, \eta^N_{s})
                - f(x)r(x, \eta^N_{s})
            \right)
        q_\theta(x,dz) \right)
%\\ & \qquad \qquad \qquad \qquad {}
         \gamma(x, \eta^N_{s}), \eta^N_{s}(dx) \bigg\rangle
\\ & \qquad \qquad \qquad \qquad \qquad {}
    + \bigg\langle f(x) F(x, \eta^N_{s}),
    \eta^N_{s}(dx) \bigg\rangle
    \bigg\} ds
\end{aligned}    
\end{equation}
is a martingale (with respect to the natural filtration), 
with quadratic variation
    \begin{equation} \label{eqn:prelimit_martingale_variation}
\begin{aligned} \relax
[ M^N(f)]_t =& 
\frac{\theta}{N} \int_{0}^{t}\bigg\{
    \Big\langle \gamma(x, \eta^N_{s})
        \int f^2(z)r(z, \eta^N_{s}) q_\theta(x,dz) 
    , \eta^N_{s}(dx) \Big\rangle \\
& \qquad \qquad {}
    + \Big\langle \mu_\theta(x, \eta^N_{s}) f^2(x) 
    , \eta^N_{s}(dx)\Big\rangle
    \bigg\}ds. 
\end{aligned}    
\end{equation}
\end{definition}
The form of~(\ref{eqn:eta_martingale}) and~(\ref{eqn:prelimit_martingale_variation})
is explained in Section~\ref{sec:heuristics}.
Note that since (juvenile) individuals are produced at rate $N \gamma \eta$, but each
has mass $1/N$,
these factors of $N$ cancel in~\eqref{eqn:eta_martingale}. 
Under our scaling, $N$ and $\theta=\theta(N)$ will tend to infinity in such a 
way that $\alpha:=\lim_{N\to\infty}\theta(N)/N$ exists and is finite.
From the expression~(\ref{eqn:prelimit_martingale_variation}) it is easy to guess that 
whether the limiting processes will be deterministic or stochastic
is determined by whether $\alpha$ is zero or nonzero.

@@

We give two concrete situations in which the fourth condition is satisfied.

@@

The purpose of the conditions that we have placed on the reproduction parameters is to 
ensure that the net per capita reproduction rate (before time scaling)
is order $1/\theta$. As remarked above, 
because of the non-local reproduction mechanism, it no longer suffices to assume that 
$r(x,\smooth{r}\eta(x))\gamma(x, \smooth{\gamma}\eta(x))-\mu_\theta(x)$ is of order $1/\theta$.
An instructive example is the case where $\gamma \equiv 1$ and $F \equiv 0$, so that $\mu_\theta = r$,
and $\eta = \delta_x$
(i.e., the population has all individuals at a single location)
in which case the mean rate of change of the total population size
is $\int (r(y, N \smooth{r}(y) ) - r(x, N \smooth{r}(x))) q_\theta(x, dy)$;
the first condition of Lemma~\ref{lem:conditions on r} would ensure this is of order $1/\theta$.

If $r(x,m)$ is independent of $m$, then the conditions are easy to 
satisfy; they just require some regularity of $r$ as a function of $x$. 
Condition~\ref{control through r} is also satisfied if for example
$\|\nabla\rho_r\|\leq C\rho_r$ 
and $m\partial_mr(x,m)$, $m^2\partial_{mm}r(x,m)$ are bounded.
This is the case, for instance, if $\rho_r$ decays exponentially.
On the other hand, it might seem more natural to take $\rho_r$ to be a Gaussian 
density with parameter $\sigma_r$, say. Then, 
as we check in Lemma~\ref{old 6.1}, \comment{this is referring to the
old Lemma 6.1, now to go in Appendix,}
Condition~\ref{control through gamma} is satisfied if $\rho_\gamma$ is also
Gaussian with parameter $\sigma_\gamma$ and $\sigma_\gamma>\sigma_r$. For large enough
$\theta$, this condition guarantees that 
$\sigma_r+1/\theta <\sigma_\gamma$, so that the establishment probability of a juvenile
is controlled by individuals that are already `felt' by the fecundity kernel $\rho_\gamma$
at the location of their parent.

@@

Our main results depend on two dichotomies:
Is the limiting process deterministic or a (generalized) superprocess?
And, are interactions pointwise in the limit or nonlocal?
Below we have results for deterministic limits with pointwise and nonlocal interactions,
and for superprocess limits with nonlocal interactions.
See Figure \ref{fig:super_vs_det_2d} for snapshots of the population
from direct simulation of the process using SLiM \citep{haller_slim_2019}.

\begin{figure}
    \begin{center}
        \includegraphics{figures/ex1a/fkpp_123.locations}
        \includegraphics{figures/ex1b/fkpp_123.locations}
    \end{center}
    \caption{
        Snapshots of two simulations, with small $\alpha=\theta/N$ (left) and large 
	$\alpha =\theta/N$ (right).
        Simulations are run with a Fisher-KPP-like parameterization:
        birth and establishment are constant, while death increases linearly with density,
        at slope $1/\theta$.
        Left: $\alpha=0.1$. Right: $\alpha=10$.
        Other parameters were the same:
        dispersal and interactions distance were set to 1,
        and the equilibrium density is 10 individuals per unit area.
        Blue curves are contours from a kernel density estimate.
        \label{fig:super_vs_det_2d}
    }
\end{figure}

\paragraph{Scaling limits with nonlocal interactions:}
Recall that the process $(\eta_t^N)_{t\geq 0}$ takes its values in the space
${\cal D}([0,\infty),\measures)$ of c\`adl\`ag
paths on $\measures$. We endow $\measures$ with the topology of weak convergence
and 
${\cal D}([0,\infty),\measures)$ with the Skorohod topology.
A sequence of processes taking values in 
${\cal D}([0,\infty),\measures)$ is said to be tight if the corresponding
sequence of distributions
is tight, so that any infinite sequence has a weakly convergent subsequence.
Our first result establishes tightness of our rescaled population processes in 
the case in which interactions remain nonlocal under the scaling, and 
characterises limit points as solutions to a martingale problem.

\begin{theorem} \label{thm:nonlocal_convergence}
    Let $(\eta^N_t)_{t \geq 0}$
    be as defined in Definition~\ref{defn:mgale_construction}
	and assume that as $N \to \infty$, $\theta(N) \to \infty$
	in such a way that $\theta(N)/N \to \alpha$.
    (However, the kernels $\rho_r$, $\rho_\gamma$, and $\rho_F$
    remain fixed.) 
Suppose that Assumptions~\ref{def:model_setup} hold and, 
further, that $\{\eta_0^N\}_{N\geq 1}$ is a sequence of 
purely atomic measures, with $\eta_0^N$ comprised of atoms of mass $1/N$, which is
tight 
in $\measures$ and such that there exists $\lambda>0$ for 
which $\langle e^{\lambda\|x\|}, \eta_0^N(dx)\rangle<C<\infty$
with $C$ independent of $N$.
    Then the sequence of processes
    $(\eta^N_t)_{t \geq 0}$ is tight, and for any limit point
    $(\eta_t)_{t \geq 0}$, for every $f \in C^\infty_b(\IR^d)$,
    \begin{align} \label{eqn:limiting_mgale_problem}
        \begin{split}
        M_t(f)
            &:=
            \langle f(x), \eta_t(dx) \rangle
            -
            \langle f(x), \eta_0(dx) \rangle
            \\ & \qquad
            -
            \int_0^t \big\langle
                \gamma(x, \eta_s)
                \mathcal{B}\left(
                    f(\cdot) r(\cdot, \eta_s)
                \right)(x)
                %\\ &\qquad \qquad \qquad {}
                +
                f(x)
                F(x, \eta_s),
                \eta_{s}(dx)
            \big\rangle ds
        \end{split}
    \end{align}
    is a martingale (with respect to the natural filtration),
    with quadratic variation
    \begin{align} \label{eqn:limiting_mgale_variation}
        [ M(f) ]_t
        =
        \alpha
        \int_0^t
        \big\langle
            %\left(
	    2\gamma\left( x, \eta_{s} \right)
            r\left(x, \eta_{s} \right)% + \mu\left(x, \eta_{s} \right)\right)
            f^2(x),
            \eta_{s} (dx)
        \big\rangle ds. 
    \end{align}
    Furthermore, if $\alpha = 0$ the limit is deterministic.
\end{theorem}

@@

\begin{corollary} \label{cor:superprocess_uniqueness}
    Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
    if the martingale problem
    defined by equations~\eqref{eqn:limiting_mgale_problem} 
and~\eqref{eqn:limiting_mgale_variation}
    has a unique solution,
    then $(\eta^N_t)_{t \ge 0}$ converges weakly
to that solution
    as $N \to \infty$.
\end{corollary}
When $\alpha>0$, the limit points can be thought of as interacting superprocesses. For
example, when $r$ and $\gamma$ are constant, and 
$F(x,\eta_s)=1-\smooth{F} \eta_s(x)$,
we recover a superprocess with nonlinear death rates corresponding to logistic growth
that is a continuous limit of the Bolker-Pacala model \citep{bolker/pacala:1997,bolker/pacala:1999}.
We are not aware of a general result to determine when we will have uniqueness of 
solutions to the martingale problem of Theorem~\ref{thm:nonlocal_convergence}
when $\alpha>0$.
However, the Dawson--Girsanov transform tells us that we have uniqueness in this special 
case of the superprocess with nonlinear death rates, and 
Perkins stochastic calculus (and its adaptation to a 
lookdown setting) provides uniqueness for cases with interactions in the 
dispersal mechanism of
the superprocess. We refer 
to~\cite{dawson:1993, perkins:1992}, and \cite{donnelly/kurtz:1999} 
for approaches to showing that the martingale problem is well-posed in the case $\alpha>0$.

@@

\begin{theorem}[Convergence to a PDE]
    \label{thm:local_convergence}
    Let $(\eta^N_t)_{t \geq 0}$
    be as defined in Definition~\ref{defn:mgale_construction} with
$r(x,m)\equiv 1\equiv \gamma(r,m)$, $F(x,m)\equiv F(m)$, 
$\rho^\epsilon_F$ a symmetric Gaussian density with variance
parameter $\epsilon^2$,
and $\DG=\Delta$. 
Further suppose that $F(m)$ is a polynomial with $F(m)\ind_{m\geq 0}$ bounded above.
We assume that $\langle 1,\eta_0\rangle$ is bounded, and that for all 
$x\in\IR^d$ and $k\in \IN$,
	\[
		\limsup_{\epsilon\to 0} 
		\IE\big[\rho_F^\epsilon*\eta_0(x)^k\big]<\infty,
	\]
	and 
\[
	\int \IE\big[\rho_F^\epsilon*\eta_0(x)^k\big] dx<\infty.
\]
Finally assume that $N \to \infty$, $\theta \to \infty$
and $\epsilon \to 0$ in such a way that
\begin{equation}
\frac{1}{\theta\epsilon^2}+\frac{\theta}{N\epsilon^d}\to 0.
\end{equation}
Then the sequence of 
${\cal D}([0,\infty), \measures)$-valued
stochastic processes $\big(\rho_F^\epsilon*\eta^N_t(x)dx\big)_{t \ge 0}$
converges weakly
to a measure-valued process with a density $\varphi(t, x)$
that solves
\begin{align}
	\label{target equation in local convergence}
        \partial_t \varphi(t, x)=\Delta \varphi (t,x) +\varphi(t,x) F(\varphi(t,x)).
\end{align}
\end{theorem}

\begin{remark}
In fact,
our proof goes through without significant change under the conditions that
$F(m)\ind_{m\geq 0}$ is bounded above (but not necessarily below), and that
for all $m,n\in[0,\infty)$
\[
|F(m)|\leq\sum_{j=1}^ka_jm^j, \quad\mbox{ and }
|F(n)-F(m)|
\leq|n-m|\sum_{j=1}^{k'}b_j\Big(n^j+m^j\Big),
\]
for some non-negative constants $\{a_j\}_{j=0}^k$, $\{b_j\}_{j=0}^{k'}$.
We take $F$ to be polynomial to somewhat simplify notation in the proof.
\end{remark}
